---
title: "Capstone Project"
author: "MK NYARKO"
date: "2025-10-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Uploading packages

library(readr)
library(dplyr) #cleaning
library(tidyverse)
library(gtsummary)
library(labelled)
library(gt)
library(forcats)
library(stringr)
library(flextable)
library(tidyr)
library(data.table)
library(lubridate)
library(stringr)
library(officer)
library(nnet) #multinomial regression
library(broom) #tidy results
library(pscl) #assessing model fit (McFadden pseudo r square)
library(ggplot2) #plots
library(rvg) #plots
library(gdtools)
```

```{r, message=FALSE, warning=FALSE}
# Loading the dataset from Github Repository
url <- "https://raw.githubusercontent.com/mkNyarko/capstonemph/refs/heads/main/semaglutide_cleanPS_capstone.csv"
semaglutide_cleanPS_capstone <- read_csv(url)
head(semaglutide_cleanPS_capstone)
```

```{r}
summary(semaglutide_cleanPS_capstone)
```

```{r}
# RE-CONVERTING ALL VARIABLES TO NUMERIC, FACTOR, CHARACTER OR DATE VARIABLES AS NEEDED

semaglutide_cleanPS_capstone2 <- semaglutide_cleanPS_capstone %>%
  mutate(
    # Identifiers
    caseid = as.character(caseid),               # unique ID, not numeric
    caseversion = as.numeric(caseversion),       # version number for updates
    
    # Demographics
    age = as.numeric(age),                       # continuous variable (years)
    age_group = factor(age_group, 
                       levels = c("YOUNGER ADULTS (18–64)", "OLDER ADULTS (≥65)")),  # categorical
    age_cod = as.character(age_cod),             # usually "YR", "DEC", etc.
    sex = factor(sex, levels = c("M", "F", "UNK")),                    # categorical (male, female, unknown)
    wt = as.numeric(wt),                         # continuous (kg)
    wt_cod = as.character(wt_cod),               # usually "KG" or similar
    
    # Reporter / country
    reporter_country = factor(reporter_country), # categorical (e.g., US)
    occr_country = factor(occr_country),         # location of occurrence
    
    # Drug info
    drugname = factor(drugname),                 # categorical (brand)
    drug_standardized = factor(drug_standardized),      # categorical (brand)
    prod_ai = as.character(prod_ai),             # active ingredient(s)
    dose_amt = as.numeric(dose_amt),             # numeric dose
    dose_unit = as.character(dose_unit),         # dose unit text (mg, ml, etc.)
    role_cod = factor(role_cod, levels = c("PS", "SS", "C", "I")),     # primary/secondary suspect roles
    
    # Event info
    pt = factor(pt),                             # MedDRA-coded preferred term (event)
    outc_cod = factor(outc_cod),                 # outcome codes (DE, HO, OT, etc.)
    event_dt = as.Date(as.character(event_dt), format = "%Y%m%d"),     # date of event
    rept_dt = as.Date(as.character(rept_dt), format = "%Y%m%d"),       # date of report
    quarter = factor(quarter, 
                     levels = c("2024Q1", "2024Q2", "2024Q3", "2024Q4"))  # time period
  )

summary(semaglutide_cleanPS_capstone2)
```

```{r}
# Dropping event and report dates

semaglutide_cleanPS_capstone2 <- semaglutide_cleanPS_capstone2 %>%
  select(-event_dt, -rept_dt)

summary(semaglutide_cleanPS_capstone2)
```

```{r}
# Cleaning out NR and Other from outc_cod

semaglutide_cleanPS_capstone2 <- semaglutide_cleanPS_capstone2 %>%
  filter(!outc_cod %in% c("NR", "Other")) %>%
  mutate(outc_cod = droplevels(outc_cod))

summary(semaglutide_cleanPS_capstone2)
```

```{r}
# Keeping RI (Required Intervention) and dropping CA (Congenital Anomaly) as CA is rare for older adults and is not relevant for this study

semaglutide_cleanPS_capstone2 <- semaglutide_cleanPS_capstone2 %>%
  filter(outc_cod %in% c("DE", "HO", "LT", "DS", "OT", "RI")) %>%
  mutate(outc_cod = droplevels(outc_cod))

summary(semaglutide_cleanPS_capstone2)
```

```{r}
# Cleaning out NA's and Unknown from Sex Variable
semaglutide_cleanPS_capstone2 <- semaglutide_cleanPS_capstone2 %>%
  filter(
    !is.na(sex),
    sex != "UNK"
  ) %>%
  mutate(
    sex = droplevels(sex)
  )

summary(semaglutide_cleanPS_capstone2)
```


```{r}
#Re-labelling outc_cod to full names variable "outc_cod_full"

semaglutide_cleanPS_capstone2 <- semaglutide_cleanPS_capstone2 %>%
  mutate(
    outc_cod_full = recode(
      outc_cod,
      "DE" = "Death",
      "HO" = "Hospitalization",
      "LT" = "Life-threatening",
      "DS" = "Disability",
      "RI" = "Required intervention",
      "OT" = "Other serious outcome",
      .default = NA_character_
    )
  ) %>%
  relocate(outc_cod_full, .after = outc_cod)
summary(semaglutide_cleanPS_capstone2)
```

```{r}
# Creating pt groupings

# Psychiatric-related PTs
psychiatric_pts <- c(
  "ANXIETY", "DEPRESSION", "DEPRESSED MOOD", "MAJOR DEPRESSION",
  "SUICIDAL IDEATION", "SUICIDE ATTEMPT", "COMPLETED SUICIDE",
  "PANIC ATTACK", "INSOMNIA", "SLEEP DISORDER", "NIGHTMARE",
  "HALLUCINATION", "HALLUCINATION, AUDITORY", "DELUSION",
  "AGITATION", "IRRITABILITY", "MOOD ALTERED", "MOOD SWINGS",
  "APATHY", "ANHEDONIA", "CONFUSIONAL STATE",
  "MEMORY IMPAIRMENT", "DISTURBANCE IN ATTENTION",
  "MENTAL DISORDER", "NERVOUSNESS", "STRESS",
  "SELF-INJURIOUS IDEATION", "PSYCHOTIC BEHAVIOUR",
  "EUPHORIC MOOD", "HYPOMANIA"
)

# Cardiovascular PTs
cardiovascular_pts <- c(
  "CARDIAC ARREST", "CARDIAC FAILURE", "CARDIAC FAILURE CONGESTIVE",
  "ATRIAL FIBRILLATION", "ARRHYTHMIA", "TACHYCARDIA",
  "BRADYCARDIA", "SINUS TACHYCARDIA",
  "MYOCARDIAL INFARCTION", "PALPITATIONS",
  "HYPERTENSION", "HYPOTENSION",
  "CEREBROVASCULAR ACCIDENT", "ISCHAEMIC STROKE",
  "EMBOLIC STROKE", "TRANSIENT ISCHAEMIC ATTACK",
  "DEEP VEIN THROMBOSIS", "PULMONARY EMBOLISM",
  "CORONARY ARTERY OCCLUSION", "HEART RATE INCREASED",
  "HEART RATE DECREASED", "SYNCOPE", "PRESYNCOPE"
)

# Gastrointestinal & Metabolic PTs
gi_metabolic_pts <- c(
  "NAUSEA", "VOMITING", "DIARRHOEA", "CONSTIPATION",
  "ABDOMINAL PAIN", "ABDOMINAL PAIN UPPER", "ABDOMINAL PAIN LOWER",
  "DYSPEPSIA", "GASTROOESOPHAGEAL REFLUX DISEASE",
  "ILEUS", "INTESTINAL OBSTRUCTION", "IMPAIRED GASTRIC EMPTYING",
  "PANCREATITIS", "PANCREATITIS ACUTE", "PANCREATITIS NECROTISING",
  "CHOLELITHIASIS", "CHOLECYSTITIS",
  "DECREASED APPETITE", "INCREASED APPETITE",
  "WEIGHT DECREASED", "ABNORMAL LOSS OF WEIGHT",
  "HYPOGLYCAEMIA", "HYPERGLYCAEMIA",
  "DIABETIC KETOACIDOSIS", "EUGLYCAEMIC DIABETIC KETOACIDOSIS",
  "BLOOD GLUCOSE INCREASED", "BLOOD GLUCOSE DECREASED",
  "HEPATIC ENZYME INCREASED", "DRUG-INDUCED LIVER INJURY",
  "MALNUTRITION", "DEHYDRATION"
)
```

```{r}
# Creating the event_category variable

semaglutide_cleanPS_capstone2 <- semaglutide_cleanPS_capstone2 %>%
  mutate(
    event_category = case_when(
      pt %in% psychiatric_pts     ~ "Psychiatric-related",
      pt %in% cardiovascular_pts  ~ "Cardiovascular",
      pt %in% gi_metabolic_pts    ~ "Gastro-Intestinal and Metabolic",
      TRUE                        ~ "Others"
    )
  )%>%
  relocate(event_category, .after = pt)
summary(semaglutide_cleanPS_capstone2)
```

```{r}
# Converting event_category to factor variable

semaglutide_cleanPS_capstone2 <- semaglutide_cleanPS_capstone2 %>%
  mutate(
    event_category = factor(event_category)                           
  )

summary(semaglutide_cleanPS_capstone2)
```

```{r}
# Cleaning labels and factor levels to prevent % errors

semaglutide_table1 <- semaglutide_cleanPS_capstone2 %>%
  mutate(
    age_group = factor(
      age_group,
      levels = c("YOUNGER ADULTS (18–64)", "OLDER ADULTS (≥65)"),
      labels = c("Younger Adults (18–64)", "Older Adults (≥65)")
    ),
    sex = factor(sex, labels = c("Male", "Female")),
    event_category = factor(
      event_category,
      levels = c(
        "Gastro-Intestinal and Metabolic",
        "Cardiovascular",
        "Psychiatric-related",
        "Others"
      )
    ),
    outc_cod_full = factor(outc_cod_full)
  )

```

```{r}
# Generating Table 1

table1 <- semaglutide_table1 %>%
  select(
    age_group,
    
    # Demographics
    sex,
    wt,
    
    # Exposure characteristics
    dose_amt,
    drug_standardized,
    quarter,
    
    # Clinical outcomes
    event_category,
    outc_cod_full
  ) %>%
  tbl_summary(
    by = age_group,
    statistic = list(
      all_categorical() ~ "{n} ({p}%)",
      all_continuous() ~ "{mean} ({sd})"
    ),
    percent = "column",
    missing = "no",
    label = list(
      sex ~ "Sex",
      wt ~ "Weight (kg)",
      dose_amt ~ "Dose Amount",
      drug_standardized ~ "Reported Drug",
      quarter ~ "Quarter of Report",
      event_category ~ "Adverse Event Category",
      outc_cod_full ~ "Outcome Category"
    )
  ) %>%
  add_overall() %>%
  bold_labels()

# Displaying the table in R Markdown
table1

```

```{r}
# Converting to publication-ready Word format

table1_flex <- as_flex_table(table1) %>%
  autofit() %>%
  bold(part = "header") %>%
  bold(j = 1) %>%        # bold variable names column
  align(align = "center", part = "all") %>%
  valign(valign = "center", part = "all")

```

```{r}
# Exporting to an editable Word document

doc <- read_docx() %>%
  body_add_par(
    "Table 1. Demographic and Clinical Characteristics of U.S. Adults Reporting Semaglutide-Associated Adverse Events, FAERS 2024",
    style = "heading 1"
  ) %>%
  body_add_flextable(table1_flex) %>%
  body_add_par(
    "Values are presented as n (%) unless otherwise indicated. Percentages are column percentages.",
    style = "Normal"
  )

print(doc, target = "Table1_Semaglutide_Capstone_Updated.docx")

```

```{r}
# Handling Missing Weight and Dose Amount Data by converting them to categorical variables instead of maintaining as continuous variable

semaglutide_cleanPS_capstone3 <- semaglutide_table1 %>%
  mutate(
    wt_cat = case_when(
      is.na(wt) ~ "Missing",
      wt < 80 ~ "<80 kg",
      wt >= 80 & wt < 100 ~ "80–99 kg",
      wt >= 100 ~ "≥100 kg"
    ),
    dose_cat = case_when(
      is.na(dose_amt) ~ "Missing",
      dose_amt < 0.5 ~ "<0.5 mg",
      dose_amt >= 0.5 & dose_amt < 1 ~ "0.5–0.99 mg",
      dose_amt >= 1 ~ "≥1 mg"
    )
  )

summary(semaglutide_cleanPS_capstone3)

```

```{r}
# The drug_standardized variable is all upper case. Converting to Title Case

semaglutide_cleanPS_capstone3 <- semaglutide_cleanPS_capstone3 %>%
  mutate(
    drug_standardized = drug_standardized %>%
      as.character() %>%
      str_to_title() %>%
      factor()
  )

# Checking
levels(semaglutide_cleanPS_capstone3$drug_standardized)

```



```{r}
# Re-ordering wt_cat and dose_cat

semaglutide_cleanPS_capstone3 <- semaglutide_cleanPS_capstone3 %>%
  relocate(wt_cat, .after = wt_cod) %>%
  relocate(dose_cat, .after = dose_unit)

names(semaglutide_cleanPS_capstone3)

```

```{r}
# Writing the new clean dataset to a CSV file

write.csv(
  semaglutide_cleanPS_capstone3,
  "semaglutide_cleanPS_capstone3.csv",
  row.names = FALSE
)
```

#New Clean Dataset for Analysis
Number of observations = 2,427
Number of variables = 23


```{r}
# Generating UPDATED Table 1 with categorical weight and dose

table1_new <- semaglutide_cleanPS_capstone3 %>%
  select(
    age_group,
    
    # Demographics
    sex,
    wt_cat,
    
    # Exposure characteristics
    dose_cat,
    drug_standardized,
    quarter,
    
    # Clinical outcomes
    event_category,
    outc_cod_full
  ) %>%
  tbl_summary(
    by = age_group,
    statistic = list(
      all_categorical() ~ "{n} ({p}%)"
    ),
    percent = "column",
    missing = "no",
    label = list(
      sex ~ "Sex",
      wt_cat ~ "Weight Category",
      dose_cat ~ "Dose Category",
      drug_standardized ~ "Reported Drug",
      quarter ~ "Quarter of Report",
      event_category ~ "Adverse Event Category",
      outc_cod_full ~ "Outcome Category"
    )
  ) %>%
  add_overall() %>%
  bold_labels()

# Display table
table1_new

```

```{r}
# Converting to publication-ready Word format

table1_new_flex <- as_flex_table(table1_new) %>%
  autofit() %>%
  bold(part = "header") %>%
  bold(j = 1) %>%        # bold variable names column
  align(align = "center", part = "all") %>%
  valign(valign = "center", part = "all")

# Exporting to an editable Word document

doc2 <- read_docx() %>%
  body_add_par(
    "Table 1. Demographic and Clinical Characteristics of U.S. Adults Reporting Semaglutide-Associated Adverse Events, FAERS 2024",
    style = "heading 1"
  ) %>%
  body_add_flextable(table1_new_flex) %>%
  body_add_par(
    "Values are presented as n (%) unless otherwise indicated. Percentages are column percentages.",
    style = "Normal"
  )

print(doc2, target = "Table1_Semaglutide_Capstone_Updated2.docx")
```

```{r}
# Generating Flow Chart for Exclusion Criteria

# Flow chart as a character vector
flow_chart <- c(
  "             Figure 1. FAERS SEMAGLUTIDE DATA CLEANING FLOW CHART",
  "",
  
  # Step 1
  " ┌──────────────────────────────────────────────────────┐",
  " | Raw data (4 quarters combined): n = 17,281           |",
  " └──────────────────────────────────────────────────────┘",
  "                     |",
  "                     v    --> Excluded: 1,310 (Duplicates)",
  " ┌──────────────────────────────────────────────────────┐",
  " | After deduplication: n = 15,971                      |",
  " └──────────────────────────────────────────────────────┘",
  "",
  
  # Step 2
  "                     |",
  "                     v    --> Excluded: 7,904 (Missing age)",
  " ┌──────────────────────────────────────────────────────┐",
  " | After cleaning age variable: n = 8,067               |",
  " └──────────────────────────────────────────────────────┘",
  "",
  
  # Step 3
  "                     |",
  "                     v    --> Excluded: 2,765 (Non-Primary Suspect)",
  " ┌──────────────────────────────────────────────────────┐",
  " | After keeping Primary Suspect (PS): n = 5,302        |",
  " └──────────────────────────────────────────────────────┘",
  "",
  
  # Step 4
  "                     |",
  "                     v    --> Excluded: 2,698 (NR / Other outcomes)",
  " ┌──────────────────────────────────────────────────────┐",
  " | After outcome cleaning: n = 2,604                    |",
  " └──────────────────────────────────────────────────────┘",
  "",
  
  # Step 5
  "                     |",
  "                     v    --> Excluded: 3 (Congenital anomaly)",
  " ┌──────────────────────────────────────────────────────┐",
  " | After excluding congenital anomaly: n = 2,601        |",
  " └──────────────────────────────────────────────────────┘",
  "",
  
  # Step 6
  "                     |",
  "                     v    --> Excluded: 174 (Missing/unknown sex)",
  " ┌──────────────────────────────────────────────────────┐",
  " | Final analytic sample: n = 2,427                     |",
  " └──────────────────────────────────────────────────────┘"
)

# Print the flow chart
cat(flow_chart, sep = "\n")
```

```{r}
# Writing flow chart to word

flowdoc <- read_docx()

for (line in flow_chart) {
  flowdoc <- body_add_par(flowdoc, line, style = "Normal")
}

print(flowdoc, target = "Fig.1 FAERS_Flow_Chart.docx")
```


ANALYSIS

# A. Disproportionality analyses (Signal Detection) using Reporting Odds Ratios (ROR) and Proportional Reporting Ratios (PRR) with 95% confidence intervals for major adverse event categories overall and by age and sex

# Question:
# "Are certain adverse event categories reported more frequently than expected with semaglutide?”

```{r}
# 1. Creating binary variable for each event category

# ROR/PRR are both based on 2x2 tables
# The event of interest must be compared to all other event

data_ror <- semaglutide_cleanPS_capstone3 %>%
  mutate(
    gi_event = ifelse(event_category == "Gastro-Intestinal and Metabolic", 1, 0),
    cardio_event = ifelse(event_category == "Cardiovascular", 1, 0),
    psych_event = ifelse(event_category == "Psychiatric-related", 1, 0)
  )
```

```{r}
# 2. Building a function to compute Reporting Odds Ratios (ROR) and Proportional Reporting Ratios (PRR)

compute_ror_prr <- function(data, exposure_var, event_var) {
  
  tab <- table(data[[exposure_var]], data[[event_var]])
  
  # Ensure 2x2 structure
  if (!all(dim(tab) == c(2, 2))) return(NULL)
  
  a <- tab[2,2]
  b <- tab[2,1]
  c <- tab[1,2]
  d <- tab[1,1]
  
  # ROR
  ror <- (a / b) / (c / d)
  se_log_ror <- sqrt(1/a + 1/b + 1/c + 1/d)
  ci_lower <- exp(log(ror) - 1.96 * se_log_ror)
  ci_upper <- exp(log(ror) + 1.96 * se_log_ror)
  
  # PRR
  prr <- (a / (a + b)) / (c / (c + d))
  
  tibble(
    ROR = ror,
    CI_lower = ci_lower,
    CI_upper = ci_upper,
    PRR = prr
  )
}
```

```{r}
# 3. Defining exposure variables

# For exposure, comparison will be done for Age variable (Younger adults vs Older adults), and Sex variable (Male vs Female)

# Converting both variables to binary variables

data_ror <- data_ror %>%
  mutate(
    age_binary = ifelse(age_group == "Older Adults (≥65)", 1, 0),
    sex_binary = ifelse(sex == "Female", 1, 0)
  )
```

```{r}
# 4. Running ROR and PRR for each event category by age

results_age <- bind_rows(
  compute_ror_prr(data_ror, "age_binary", "gi_event") %>% mutate(Event = "GI"),
  compute_ror_prr(data_ror, "age_binary", "cardio_event") %>% mutate(Event = "Cardiovascular"),
  compute_ror_prr(data_ror, "age_binary", "psych_event") %>% mutate(Event = "Psychiatric")
)

# label appropriately and re-arrange columns to make it Table 2 ready
results_age <- results_age %>%
  mutate(Group = "Age") %>%
  select(Group, Event, ROR, CI_lower, CI_upper, PRR)
```

```{r}
# 5. Running ROR and PRR for each event category by sex

results_sex <- bind_rows(
  compute_ror_prr(data_ror, "sex_binary", "gi_event") %>% mutate(Event = "GI"),
  compute_ror_prr(data_ror, "sex_binary", "cardio_event") %>% mutate(Event = "Cardiovascular"),
  compute_ror_prr(data_ror, "sex_binary", "psych_event") %>% mutate(Event = "Psychiatric")
)

# label appropriately and re-arrange columns to make it Table 2 ready
results_sex <- results_sex %>%
  mutate(Group = "Sex") %>%
  select(Group, Event, ROR, CI_lower, CI_upper, PRR)
```

```{r}
# 6. Combining the results

final_results <- bind_rows(results_age, results_sex)

final_results
```

```{r}
# 7. Formatting for interpretation

final_results2 <- final_results %>%
  mutate(
    ROR = round(ROR, 2),
    CI = paste0("(", round(CI_lower, 2), "–", round(CI_upper, 2), ")"),
    PRR = round(PRR, 2)
  ) %>%
  select(Group, Event, ROR, CI, CI_lower, CI_upper, PRR)

final_results2
```
# Interpretation
Compared with younger adults, older adults demonstrated significantly higher reporting odds of cardiovascular adverse events (ROR = 1.54, 95% CI: 1.03–2.29; PRR = 1.51)

In contrast, gastrointestinal events (ROR = 0.71, 95% CI: 0.59–0.85) and psychiatric-related events (ROR = 0.40, 95% CI: 0.26–0.62) were reported less frequently among older adults relative to younger adults

When stratified by sex, no statistically significant differences were observed in the reporting of adverse event categories between males and females, as the confidence intervals for gastrointestinal, cardiovascular, and psychiatric-related events included the null value of 1

Overall, these findings suggest potential age-related differences in adverse event reporting patterns associated with semaglutide, particularly for cardiovascular events, while sex-based differences were not evident

```{r}
# 8. Exporting final results as table 2 in word format

# Assigning final results to table 2
table2 <- final_results2 %>%
  select(Group, Event, ROR, CI, PRR) %>%
  mutate(
    ROR = round(ROR, 2),
    PRR = round(PRR, 2)
  )


# Converting table to a Word-friendly format
table2_ft <- flextable(table2) %>%
  bold(part = "header") %>%
  autofit()

# Creating Word document
doc <- read_docx()

doc <- doc %>%
  body_add_par("Table 2. Disproportionality Analysis of Adverse Event Categories by Age and Sex", 
               style = "heading 1") %>%
  body_add_flextable(table2_ft)

# Exporting to word
print(doc, target = "Table2_ROR_PRR_results.docx")

# Display table
table2_ft
```

# B. Multinomial logistic regression models assessing associations between age group and adverse outcome categories, adjusting for sex, quarter of report, dose category, and weight category.

# Question:
# "After adjusting for other factors, is age still associated with different adverse event categories?"

```{r}
# Keeping "Others" fixed as reference

semaglutide_cleanPS_capstone3$event_category <- factor(
  semaglutide_cleanPS_capstone3$event_category,
  levels = c(
    "Others",
    "Gastro-Intestinal and Metabolic",
    "Cardiovascular",
    "Psychiatric-related"
  )
)
```


```{r}
# 1. Ensuring variables for the model are factor variables

data_model <- semaglutide_cleanPS_capstone3 %>%
  mutate(
    event_category = as.factor(event_category),
    age_group = as.factor(age_group),
    sex = as.factor(sex),
    quarter = as.factor(quarter),
    dose_cat = as.factor(dose_cat),
    wt_cat = as.factor(wt_cat)
  )
```

```{r}
# 2. Setting reference categories

# Outcome reference: Other
# Age reference: Younger Adults
# Sex reference: Male

data_model <- data_model %>%
  mutate(
    event_category = relevel(event_category, ref = "Others"),
    age_group = relevel(age_group, ref = "Younger Adults (18–64)"),
    sex = relevel(sex, ref = "Male")
  )
```

Assumptions Testing

5 Assumptions

- Independent observations: Assumption satisfied by deduplication step during data cleaning.
- Predictors should not be strongly correlated with each other: Predictors (age_group, sex, quarter, dose_cat, wt_cat) are conceptually different variables. Severe multicollinearity unlikely.
- Adequate sample size (about 10-20) per outcome category: All outcome categories have above 20 observations. Assumption satisfied. 
- Separation. One predictor must not perfectly predict the outcome: Co-efficients are within expected ranges and standard errors are moderate across predictors, indicating no evidence of model separation or instability (see model stability block). Assumption satisfied.
- Model convergence. The algorithm successfully estimates parameters (as seen below). Assumption met.

```{r}
# 3. Fitting the Multinomial logistic regression model

multinom_model <- multinom(
  event_category ~ age_group + sex + quarter + dose_cat + wt_cat,
  data = data_model
)

summary(multinom_model)
```
Co-efficients and standard errors provided but not yet Odds Ratios
Co-efficients and standard errors are not huge.

```{r}
# 4. Converting co-efficients to odds ratios

model_results <- tidy(multinom_model) %>%
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  )

model_results
```

```{r}
# 5. Cleaning results for reporting

model_results_clean <- model_results %>%
  select(y.level, term, OR, CI_lower, CI_upper, p.value) %>%
  mutate(
    OR = round(OR, 2),
    CI_lower = round(CI_lower, 2),
    CI_upper = round(CI_upper, 2),
    CI = paste0("(", CI_lower, "–", CI_upper, ")")
  )

model_results_clean
```

```{r}
# 6. Making results easier to interpret by renaming columns

model_results_clean <- model_results_clean %>%
  rename(
    Outcome = y.level,
    Variable = term
  )

model_results_clean
```

```{r}
# 7. Focusing on main variable of interest (Age group)

age_results <- model_results_clean %>%
  filter(grepl("age_group", Variable))

age_results
```
# Interpretation

Gastro-Intestinal and Metabolic
Older adults had 29% lower odds of reporting gastrointestinal and metabolic adverse events compared with younger adults, relative to other types of adverse events
The association is statistically significant as the confidence interval does not include 1
Age group is independently associated with GI events

Cardiovascular
There was no statistically significant difference in the odds of cardiovascular adverse events between older and younger adults after adjusting for the other variables in the model.
Although the odds ratio is slightly above 1, the confidence interval includes 1, meaning the difference could be due to chance.

Psychiatric-related
Older adults had 53% lower odds of reporting psychiatric-related adverse events compared with younger adults, relative to other adverse event types.
The association is statistically significant as the confidence interval does not include 1
Age group is independently associated with Psychiatric-related events

```{r}
# 8. Checking model stability

summary(multinom_model)$coefficients
summary(multinom_model)$standard.errors
```
# Interpretation:
The multinomial regression model appears statistically stable, with coefficients within expected ranges and moderate standard errors across predictors, indicating no evidence of model separation or instability.

```{r}
# 9. Assessing model fit

pR2(multinom_model)
```
#Interpretation

Model fit statistics indicated that the multinomial logistic regression model improved prediction of adverse event categories compared with the null model (G² = 143.8). 
The model converged successfully. 
The McFadden pseudo-R² was 0.032, suggesting modest but acceptable explanatory power.

```{r}
# 10. Exporting results for Table 3

table3_ft <- flextable(age_results) %>%
  bold(part = "header") %>%
  autofit() %>%
  width(j = 1, width = 1.2) %>%
  width(j = 2, width = 2.5) %>%
  width(j = 3:6, width = 1)

doc <- read_docx() %>%
  body_add_par("Table 3. Multinomial Logistic Regression Results", style = "heading 1") %>%
  body_add_flextable(table3_ft)

print(doc, target = "Table3_multinomial_results.docx")

# Display table
table3_ft
```

# C. Assessment of interaction effects between age group and adverse event category.
# Question:
# "Does the association between age group and adverse event category differ by sex?"

```{r}
# 1. Confirming that the outcome is set correctly

data_model$event_category <- factor(
  data_model$event_category,
  levels = c("Others",
             "Gastro-Intestinal and Metabolic",
             "Cardiovascular",
             "Psychiatric-related")
)
```

```{r}
# 2. Fitting the interaction model

interaction_model <- multinom(
  event_category ~ age_group + sex + quarter + dose_cat + wt_cat +
                   age_group:sex,
  data = data_model
)
```

```{r}
# 3. Comparing models

# Base model vs Interaction model

anova(multinom_model, interaction_model, test = "Chisq")
```
# Interpretation

Small likelihood ratio statistic of 0.78 indicating that the model barely changed after adding interaction term
P-value >0.05. Interaction is not statistically significant.

The relationship between age group and adverse event category does not differ by sex.
Older vs younger adults show similar adverse event patterns in both males and females.


# D. Sensitivity analyses evaluating robustness of findings across alternative categorizations and exclusion of select missing-data strata

# Question
# "If slight changes are made to how variables are defined or certain uncertain data are removed, do the results stay similar?"

```{r}
# 1. Creating dataset excluding missing dose

sens_no_missing_dose <- semaglutide_cleanPS_capstone3 %>%
  filter(dose_cat != "Missing")

# Checking sample size
nrow(sens_no_missing_dose)
```

```{r}
# 2. Re-running Multinomial regression (Dose Sensitivity)

model_sens_dose <- nnet::multinom(
  event_category ~ age_group + sex + quarter + dose_cat + wt_cat,
  data = sens_no_missing_dose
)

# Extracting results
sens_dose_results <- broom::tidy(model_sens_dose, exponentiate = TRUE, conf.int = TRUE)

# Renaming columns
sens_dose_results <- sens_dose_results %>%
  rename(
    Outcome = y.level,
    Variable = term,
    OR = estimate
  )


sens_dose_results
```

```{r}
# 3. Creating dataset excluding missing weight

sens_no_missing_weight <- semaglutide_cleanPS_capstone3 %>%
  filter(wt_cat != "Missing")

# Checking sample size
nrow(sens_no_missing_weight)
```

```{r}
# 4. Re-running Multinomial regression (Weight sensitivity)

model_sens_weight <- nnet::multinom(
  event_category ~ age_group + sex + quarter + dose_cat + wt_cat,
  data = sens_no_missing_weight
)

# Extracting results
sens_weight_results <- broom::tidy(model_sens_weight, exponentiate = TRUE, conf.int = TRUE)

# Renaming columns
sens_weight_results <- sens_weight_results %>%
  rename(
    Outcome = y.level,
    Variable = term,
    OR = estimate
  )

sens_weight_results
```

```{r}
# 5. Creating dataset excluding both missing dose and weight

sens_complete_cases <- semaglutide_cleanPS_capstone3 %>%
  filter(
    dose_cat != "Missing",
    wt_cat != "Missing"
  )

# Checking sample size
nrow(sens_complete_cases)
```

```{r}
# 6. Re-running Multinomial regression (Weight + Dose sensitivity)

model_sens_complete <- nnet::multinom(
  event_category ~ age_group + sex + quarter + dose_cat + wt_cat,
  data = sens_complete_cases
)

# Extracting results
sens_complete_results <- broom::tidy(model_sens_complete, exponentiate = TRUE, conf.int = TRUE)

# Renaming columns
sens_complete_results <- sens_complete_results %>%
  rename(
    Outcome = y.level,
    Variable = term,
    OR = estimate
  )

sens_complete_results
```

```{r}
# 7. Alternative Dose Categorization

semaglutide_alt_dose <- semaglutide_cleanPS_capstone3 %>%
  mutate(
    dose_cat_alt = case_when(
      dose_amt < 0.5 ~ "Low dose",
      dose_amt >= 0.5 & dose_amt < 1 ~ "Medium dose",
      dose_amt >= 1 ~ "High dose",
      TRUE ~ "Missing"
    )
  )
```

```{r}
# Model
model_alt_dose <- nnet::multinom(
  event_category ~ age_group + sex + quarter + dose_cat_alt + wt_cat,
  data = semaglutide_alt_dose
)

model_alt_dose
```

```{r}
# 8. Comparing results of Weight + Dose sensitivity with main model

main_results <- age_results %>%
  filter(str_detect(Variable, "age_group"))

sens_results <- sens_complete_results %>%
  filter(str_detect(Variable, "age_group"))

main_results
sens_results
```
OR values vary between the 2 models
No statistically significant OR reported in Weight + Dose sensitivity test

```{r}
# 9. Comparing results of Weight sensitivity with main model

main_results <- age_results %>%
  filter(str_detect(Variable, "age_group"))

sens_results <- sens_weight_results %>%
  filter(str_detect(Variable, "age_group"))

main_results
sens_results
```
OR values vary between the 2 models, except for psychiatry-related events
Statistically significant OR reported in Weight sensitivity test for psychiatry-related events only


```{r}
# 10. Comparing results of Dose sensitivity with main model

main_results <- age_results %>%
  filter(str_detect(Variable, "age_group"))

sens_results <- sens_dose_results %>%
  filter(str_detect(Variable, "age_group"))

main_results
sens_results
```
OR values vary between the 2 models, except for Gastro-Intestinal and Metabolic events
Statistically significant OR reported in Dose sensitivity test for Gastro-Intestinal and Metabolic events only


```{r}
# Summary table for sensitivity analysis

sensitivity_summary <- bind_rows(
  sens_dose_results %>% mutate(Model = "Exclude missing dose"),
  sens_weight_results %>% mutate(Model = "Exclude missing weight"),
  sens_complete_results %>% mutate(Model = "Complete cases")
) %>%
  filter(str_detect(Variable, "age_group"))

sensitivity_summary
```
Summary Gastro-Intestinal and Metabolic OR = 0.66 - 1.26 for all 3 sensitivity models, OR main model = 0.71 
Summary Cardiovascular OR = 0.83 - 1.30 for all 3 sensitivity models, OR main model = 1.15
Summary Psychiatric-related OR = 0.13 - 0.66 , OR main model = 0.47

The main model OR values fall within range of the OR values obtained from the  sensitivity models.

Results are stable. Model is robust.



# E. GGPLOT2 FIGURES

```{r}
# 1. Generating fig.2 Forest Plot of disproportionality signals

plot1_data <- final_results2 %>%
  filter(Group == "Age") %>%
  mutate(Event = factor(Event, levels = rev(unique(Event))))

plot1 <- ggplot(plot1_data, aes(x = ROR, y = Event)) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = CI_lower, xmax = CI_upper), height = 0.2) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  labs(
    title = "Figure 2. Reporting Odds Ratios for Major Adverse Event Categories",
    x = "Reporting Odds Ratio (95% CI)",
    y = "Adverse Event Category"
  ) +
  theme_minimal()

plot1
```

```{r}
# Exporting fig 2 as png file

ggsave("Figure2_ROR_Forest_Plot.png", plot1, width = 9, height = 4, dpi = 600)
```

```{r}
# 2. Generating fig.3 Sub-group Forest Plot (Age and Sex)

plot2_data <- final_results2 %>%
  mutate(label = paste(Group, Event, sep = ": "),
         label = factor(label, levels = rev(unique(label))))

plot2 <- ggplot(plot2_data, aes(x = ROR, y = label)) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = CI_lower, xmax = CI_upper), height = 0.2) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  labs(
    title = "Figure 3. Disproportionality Signals by Age and Sex",
    x = "Reporting Odds Ratio (95% CI)",
    y = "Subgroup and Event"
  ) +
  theme_minimal()

plot2
```

```{r}
# Exporting fig 3 as png file

ggsave("Figure3_Subgroup_Forest_Plot.png", plot2, width = 9, height = 4, dpi = 600)
```

```{r}
# 3. Generating fig.4 Distribution of adverse events

plot3 <- ggplot(event_distribution, aes(x = event_category, y = n, fill = event_category)) +
  geom_bar(stat = "identity", width = 0.7) +
  labs(
    title = "Figure 4. Distribution of Adverse Event Categories",
    x = "Adverse Event Category",
    y = "Number of Reports",
    fill = "Event Category"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "right",
    axis.text.x = element_text(angle = 20, hjust = 1)
  )

plot3
```

```{r}
# Exporting fig 4 as png file
ggsave(
  "Figure4_Event_Distribution.png",
  plot3,
  width = 8,
  height = 5,
  dpi = 600
)
```






















