---
title: "Capstone Project"
author: "MK NYARKO"
date: "2025-10-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(gtsummary)
library(labelled)

```

```{r, message=FALSE, warning=FALSE}
# Loading the dataset from Github Repository
url <- "https://raw.githubusercontent.com/mkNyarko/capstonemph/refs/heads/main/semaglutide_cleanPS_capstone.csv"
semaglutide_cleanPS_capstone <- read_csv(url)
head(data)
```

```{r}
summary(semaglutide_cleanPS_capstone)
```

```{r}
# RE-CONVERTING ALL VARIABLES TO NUMERIC, FACTOR, CHARACTER OR DATE VARIABLES AS NEEDED

semaglutide_cleanPS_capstone <- semaglutide_cleanPS_capstone %>%
  mutate(
    # Identifiers
    caseid = as.character(caseid),               # unique ID, not numeric
    caseversion = as.numeric(caseversion),       # version number for updates
    
    # Demographics
    age = as.numeric(age),                       # continuous variable (years)
    age_group = factor(age_group, 
                       levels = c("YOUNGER ADULTS (18–64)", "OLDER ADULTS (≥65)")),  # categorical
    age_cod = as.character(age_cod),             # usually "YR", "DEC", etc.
    sex = factor(sex, levels = c("M", "F", "UNK")),                    # categorical (male, female, unknown)
    wt = as.numeric(wt),                         # continuous (kg)
    wt_cod = as.character(wt_cod),               # usually "KG" or similar
    
    # Reporter / country
    reporter_country = factor(reporter_country), # categorical (e.g., US)
    occr_country = factor(occr_country),         # location of occurrence
    
    # Drug info
    drugname = factor(drugname),                 # categorical (brand)
    drug_standardized = factor(drug_standardized),      # categorical (brand)
    prod_ai = as.character(prod_ai),             # active ingredient(s)
    dose_amt = as.numeric(dose_amt),             # numeric dose
    dose_unit = as.character(dose_unit),         # dose unit text (mg, ml, etc.)
    role_cod = factor(role_cod, levels = c("PS", "SS", "C", "I")),     # primary/secondary suspect roles
    
    # Event info
    pt = factor(pt),                             # MedDRA-coded preferred term (event)
    outc_cod = factor(outc_cod),                 # outcome codes (DE, HO, OT, etc.)
    event_dt = as.Date(as.character(event_dt), format = "%Y%m%d"),     # date of event
    rept_dt = as.Date(as.character(rept_dt), format = "%Y%m%d"),       # date of report
    quarter = factor(quarter, 
                     levels = c("2024Q1", "2024Q2", "2024Q3", "2024Q4"))  # time period
  )

summary(semaglutide_cleanPS_capstone)
```

```{r descriptives-table, message=FALSE, warning=FALSE}

# If not already loaded:
# data <- read_csv("semaglutide_cleanPS_capstone.csv")

# Add readable labels for variables
var_label(data$age_group) <- "Age group"
var_label(data$sex)       <- "Sex"
var_label(data$wt)        <- "Weight (kg)"
var_label(data$dose_amt)  <- "Dose amount"
var_label(data$drugname)  <- "Drug name"
var_label(data$outc_cod)  <- "Outcome code"
var_label(data$quarter)   <- "Quarter of report"
var_label(data$pt)        <- "Preferred term (Adverse event type)"

# Create descriptive summary table grouped by age group
table1 <- data %>%
  select(age_group, sex, wt, dose_amt, drugname, pt, outc_cod, quarter) %>%
  tbl_summary(
    by = age_group,
    label = list(
      sex ~ "Sex",
      wt ~ "Weight (kg)",
      dose_amt ~ "Dose amount",
      drugname ~ "Reported drug",
      pt ~ "Preferred term (Adverse event type)",
      outc_cod ~ "Outcome category",
      quarter ~ "Quarter of report"
    ),
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "no"
  ) %>%
  modify_header(label ~ "**Variable**") %>%
  modify_caption("**Table 1. Descriptive Characteristics of Reported Semaglutide Adverse Event Cases by Age Group, 2024 (FAERS Data)**") %>%
  bold_labels()

# Display the table nicely in R Markdown
table1
































